hbm_paper_results

### 2024-06-14 paper_lbcc_results.Rmd
changes in extreme centile categorization between combat-gam and combatls

mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
```

@@

>>>>

mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)

summary(mean_cent_deltas_df_study$abs_mean_delta_cent)
```

see how often features move between extreme (>95% or <5%) to not-extreme
```{r}
#rm features that didn't converge in both datasets
pheno_list_cent <- pheno_list[pheno_list != "lh_SA_posteriorcingulate" & pheno_list != "rh_CT_posteriorcingulate"]

gam_centile_df.clean$combat <- "cf.gam"
gamlss_centile_df.clean$combat <- "cf.gamlss"

centiles_df <- full_join(gam_centile_df.clean, gamlss_centile_df.clean)

centiles_cat <- centiles_df %>%
  mutate(combat = factor(combat, ordered=TRUE, levels=c("cf.gam", "cf.gamlss"))) %>%
  mutate(across(pheno_list_cent, ~ as.factor(case_when(. > 0.95 ~ "high", 
                                               . < 0.05 ~ "low",
                                               TRUE ~ "mid")))) %>%
  arrange(combat) %>%
  group_by(participant) %>%
  mutate(across(pheno_list_cent, 
                ~ . == first(.), .names = "is_equal_{col}")) %>%
  ungroup()
```

```{r}
centiles_cat_filt <- centiles_cat %>%
  dplyr::select(participant, combat, all_of(pheno_list_cent))

ext_count_df <- centiles_cat_filt %>%
  rowwise() %>%
  mutate(count_high = sum(c_across(pheno_list_cent) == "high"),
         count_mid = sum(c_across(pheno_list_cent) == "mid"),
         count_low = sum(c_across(pheno_list_cent) == "low")) %>%
  dplyr::select(participant, combat, starts_with("count"))

ext_count_df %>%
  ggplot() +
  geom_bar(aes(x=count_high, fill=combat, color=combat), position="identity", alpha=.4)
ext_count_df %>%
  ggplot() +
  geom_bar(aes(x=count_low, fill=combat, color=combat), position="identity", alpha=.4)
```

```{r}
centiles_cat_filt2 <- centiles_cat %>%
  dplyr::filter(combat =="cf.gamlss") %>%
  dplyr::select(participant, starts_with("is_equal_")) %>%
  mutate(sum = rowSums(across(starts_with("is_equal_")))) %>%
  mutate(changed = ifelse(sum == length(pheno_list_cent), "no_change", "change")) #if not all features are true -> changed

max(centiles_cat_filt2$sum) == length(pheno_list_cent) #correct

length(pheno_list_cent) - mean(centiles_cat_filt2$sum) #mean per subject
length(pheno_list_cent) - min(centiles_cat_filt2$sum) #range
length(pheno_list_cent) - max(centiles_cat_filt2$sum)

#percentage subjects
table(centiles_cat_filt2$changed)["change"]/nrow(centiles_cat_filt2)

#percentage features
#n changed features = # total features - sum
n_feat <- n_unique(centiles_cat_filt2$participant)*length(pheno_list_cent)
n_unchanged <- sum(centiles_cat_filt2$sum) #true = 1 = no change

n_feat-n_unchanged
((n_feat-n_unchanged)/n_feat)*100 # % changed
```

plot
```{r}
centiles_cat_filt2_long <- centiles_cat_filt2 %>%
  pivot_longer(cols=starts_with("is_equal_"), names_to="pheno", names_prefix="is_equal_", values_to="is_equal")

n_unique(centiles_cat_filt2_long$participant)*206 == nrow(centiles_cat_filt2_long)

cent_cat_plt <- centiles_cat_filt2_long %>%
  dplyr::filter(is_equal==FALSE) %>%
ggplot() +
  geom_bar(aes(x=pheno), fill="#1D213B") +
  theme_linedraw()+
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) +
  labs(x="Brain Feature", y= "Instances of Centile Re-categorization")
ggsave("figs/cent_cat_plt.jpeg", cent_cat_plt, width=10, height=6, units="in")
```

@@

```{r}
raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info) %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)
```
